HTW Model e1

### 2024-03-21 Model/htw_model.qmd
tweaking

---
title: HTW Modeling 
author: Thomas Gorman
date:  last-modified #"`r Sys.Date()`"
page-layout: full
categories: [Modeling, ALM, EXAM, R]
lightbox: true
toc: false
code-fold: true
code-tools: true
execute: 
  warning: false
  eval: true
format:
  html: default
  hugo-md:
    echo: false
    html-math-method: mathjax
    output-file: e1-hugo.md
  gfm:
    echo: true
    output-file: e1-gfm.md
  docusaurus-md:
    echo: true
    output-file: e1-docusaurus.md
#prefer-html: false
---

```{r}
pacman::p_load(dplyr,purrr,tidyr,ggplot2, data.table, here, patchwork, conflicted, 
               stringr,future,furrr, knitr, reactable, flextable,ggstance, htmltools,ggdist)
#conflict_prefer_all("dplyr", quiet = TRUE)
walk(c("flextable","dplyr"), conflict_prefer_all, quiet = TRUE)
#options(brms.backend="cmdstanr",mc.cores=4)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("Display_Functions","fun_alm","fun_indv_fit","fun_model", "prep_model_data"), ~source(here::here(paste0("Functions/", .x, ".R"))))

@@

```{r}
#| label: fig-htw-resid-pred
#| fig-cap:  Model residuals for each combination of training condition, fit method, and model. Residuals reflect the difference between observed and predicted values. Lower values indicate better model fit. Note that y axes are scaled differently between facets. 
#| fig-height: 14
#| fig-width: 11
##| layout: [[45,-5, 45], [100]]
##| fig-subcap: ["Model Residuals - training data", "Model Residuals - testing data","Full posterior predictive distributions vs. observed data from participants."]
train_resid <- pd_train |> group_by(id,condit,Model,Fit_Method, Block) |> 
  summarise(y = mean(y), pred = mean(pred), error = y - pred) |>
  ggplot(aes(x = Model, y = abs(error), fill=factor(Block))) + 
  stat_bar + 
  ggh4x::facet_nested_wrap(rename_fm(Fit_Method)~condit, scales="free",ncol=2) +
  scale_fill_manual(values=wes_palette("AsteroidCity2"))+
  labs(title="Model Residual Errors - Training Stage", y="RMSE", x= "Model") +
  theme(legend.title = element_blank(), legend.position="top")

test_resid <-  post_dat |> 
   group_by(id,condit,x,Model,Fit_Method,rank) |>
   summarize(error=mean(abs(y-pred)),n=n()) |>
   group_by(id,condit,x,Model,Fit_Method) |>
   summarize(error=mean(error)) |>
  mutate(vbLab = factor(paste0(x,"-",x+200))) |>
  ggplot(aes(x = Model, y = abs(error), fill=vbLab)) + 
  stat_bar + 
  scale_fill_manual(values=wes_palette("AsteroidCity2"))+
  ggh4x::facet_nested_wrap(rename_fm(Fit_Method)~condit, axes = "all",ncol=2,scale="free") +
  labs(title="Model Residual Errors - Testing Stage",y="RMSE", x="Velocity Band") +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5))

@@

The posterior distributions of the $c$ and $lr$ parameters are shown @fig-htw-post-dist, and model predictions are shown alongside the empirical data in fig-cm-vx-pat and fig-cm-dev-pat (i.e. these plots combine all the posterior samples from all of the subjects). There were substantial individual differences in the posteriors of both parameters, with the within-group individual differences generally swamped any between-group or between-model differences. The magnitude of these individual differences remains even if we consider only the single best parameter set for each subject.

@@

```{r}
#| label: fig-cm-vx-pat
#| fig-height: 11
#| fig-width: 11

@@

pemp1 <- e1 |> filter(expMode2=="Test") |> mutate(Resp="Observed") |> 
  ggplot(aes(x=condit,y=vx, fill=vb)) + 
  stat_bar+ 
  labs(y="vx", x="Condition") + ggh4x::facet_nested_wrap(~Resp)

layout <- "
#A#
#B#
CCC
"

pmod1 <- post_dat_l |> filter(!(Resp=="Observed")) |> 
  group_by(id,condit, Fit_Method,Resp,x) |> 
  mutate(x=as.factor(x)) |>
 summarize(vx=mean(val)) |> 
  ggplot(aes(x=condit,y=vx, fill=x)) + 
  stat_bar + 
    ggh4x::facet_nested_wrap(~Resp+rename_fm(Fit_Method), axes = "all",ncol=3,scale="free") + 
  theme(legend.position = "none",plot.title = element_text(hjust=.55) ) + labs(title="Model Predictions", y="vx", x="Condition")

(pemp1) / plot_spacer() / pmod1 + 
  plot_layout(design = layout, heights = unit(c(5,-5, 8), c('cm','null')))

@@

```{r}
#| label: fig-cm-dev-pat
#| fig-height: 11
#| fig-width: 11
#| fig-cap: "Posterior predictives for absolute deviation from target (accuracy). Observed data is shown on top, and predictions from each combination of model and fitting method are shown below. "

pemp1 <- e1 |> filter(expMode2=="Test") |> mutate(Resp="Observed") |> 
  ggplot(aes(x=condit,y=dist, fill=vb)) + 
  stat_bar+ theme(legend.position="top")
  labs(y="Absolute Deviation", x="Condition") + ggh4x::facet_nested_wrap(~Resp)

layout <- "
#A#
#B#
CCC
"

pmod1 <- post_dat_l |> filter(!(Resp=="Observed")) |> 
  group_by(id,condit, Fit_Method,Resp,x) |> 
  mutate(x=as.factor(x)) |>
 summarize(dist=mean(dist)) |> 
  ggplot(aes(x=condit,y=dist, fill=x)) + 
  stat_bar + 
    ggh4x::facet_nested_wrap(~Resp+rename_fm(Fit_Method), axes = "all",ncol=3) + 
  theme(legend.position = "none",plot.title = element_text(hjust=.55) ) + labs(title="Model Predictions", y="Absolute Deviation", x="Condition")

# (pemp1) / plot_spacer() / pmod1 + 
#   plot_layout(design = layout, heights = unit(c(4,-5, 9), c('cm','null')))

(pemp1) / plot_spacer() / pmod1 + 
  plot_layout(design = layout, heights = unit(c(5,-5, 8), c('cm','null')))

# (pemp1) / pmod1 + plot_layout(design = layout,heights = unit(c(2,3),c("null","null")))
# 
# 
# (pemp1) / plot_spacer() / pmod1 + 
#   plot_layout(design = layout, heights = unit(c(6,-5, 8), c('cm','null')))

@@

@@

>>>>

---
title: HTW Modeling 
author: Thomas Gorman
date:  last-modified #"`r Sys.Date()`"
page-layout: full
categories: [Modeling, ALM, EXAM, R]
lightbox: true
toc: false
code-fold: true
code-tools: true
execute: 
  warning: false
  eval: true
format:
  html: default
  hugo-md:
    echo: false
    html-math-method: mathjax
    output-file: e1-hugo.md
  gfm:
    echo: true
    output-file: e1-gfm.md
  docusaurus-md:
    echo: true
    output-file: e1-docusaurus.md
prefer-html: true
---

```{r}
pacman::p_load(dplyr,purrr,tidyr,ggplot2, data.table, here, patchwork, conflicted, 
               stringr,future,furrr, knitr, reactable, flextable,ggstance, htmltools,ggdist,ggh4x)
#conflict_prefer_all("dplyr", quiet = TRUE)
walk(c("flextable","dplyr"), conflict_prefer_all, quiet = TRUE)
#options(brms.backend="cmdstanr",mc.cores=4)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("Display_Functions","fun_alm","fun_indv_fit","fun_model", "prep_model_data"), ~source(here::here(paste0("Functions/", .x, ".R"))))

@@

```{r}
#| label: fig-htw-resid-pred
#| fig-cap:  Model residuals for each combination of training condition, fit method, and model. Residuals reflect the difference between observed and predicted values. Lower values indicate better model fit. Note that y axes are scaled differently between facets. 
#| fig-height: 14
#| fig-width: 11
##| layout: [[45,-5, 45], [100]]
##| fig-subcap: ["Model Residuals - training data", "Model Residuals - testing data","Full posterior predictive distributions vs. observed data from participants."]
train_resid <- pd_train |> group_by(id,condit,Model,Fit_Method, Block,x) |> 
  summarise(y=mean(y), pred=mean(pred), mean_error=abs(y-pred)) |>
  group_by(id,condit,Model,Fit_Method,Block) |>
  summarise(mean_error=mean(mean_error)) |>
  ggplot(aes(x=interaction(Block,Model), y = mean_error, fill=factor(Block))) + 
  stat_bar + 
  ggh4x::facet_nested_wrap(rename_fm(Fit_Method)~condit, scales="free",ncol=2) +
   scale_x_discrete(guide = "axis_nested") +
  scale_fill_manual(values=c("gray10","gray50","gray92"))+
  labs(title="Model Residual Errors - Training Stage", y="RMSE", x= "Model",fill="Training Block") +
  theme(legend.position="top")

test_resid <-  post_dat |> 
   group_by(id,condit,x,Model,Fit_Method) |>
    summarise(y=mean(y), pred=mean(pred), error=abs(y-pred)) |> 
  mutate(vbLab = factor(paste0(x,"-",x+200))) |>
  ggplot(aes(x = Model, y = abs(error), fill=vbLab)) + 
  stat_bar + 
  #scale_fill_manual(values=wes_palette("AsteroidCity2"))+
  ggh4x::facet_nested_wrap(rename_fm(Fit_Method)~condit, axes = "all",ncol=2,scale="free") +
  labs(title="Model Residual Errors - Testing Stage",y="RMSE", x="Velocity Band")

@@

The posterior distributions of the $c$ and $lr$ parameters are shown @fig-htw-post-dist, and model predictions are shown alongside the empirical data in @fig-cm-vx-pat and @fig-cm-dev-pat (i.e. these plots combine all the posterior samples from all of the subjects). There were substantial individual differences in the posteriors of both parameters, with the within-group individual differences generally swamped any between-group or between-model differences. The magnitude of these individual differences remains even if we consider only the single best parameter set for each subject.

@@

```{r}
#| label: fig-cm-vx-pat
#| fig-height: 12
#| fig-width: 11

@@

pemp1 <- e1 |> filter(expMode2=="Test") |> mutate(Resp="Observed") |> 
  ggplot(aes(x=condit,y=vx, fill=vb, col=ifelse(bandType=="Trained","black",NA),size=ifelse(bandType=="Trained","black",NA))) + 
  stat_bar+ 
  scale_color_manual(values = c("black" = "black"), guide = FALSE) +
  scale_size_manual(values = c("black" = .5), guide = FALSE) + 
  theme(legend.position="right", axis.title.x = element_blank(), plot.title = element_text(hjust=.50)) +
  labs(title="Empirical Data - Experiment 1",y="vx", x="Condition",fill="Band")

layout <- "
#A#
CCC
"

# pmod1 <- post_dat_l |> filter(!(Resp=="Observed")) |> 
#   group_by(id,condit, Fit_Method,Resp,x) |> 
#  summarize(vx=mean(val)) |> 
#  left_join(testAvgE1, by=join_by(id,condit,x==bandInt)) |>
#   ggplot(aes(x=condit,y=vx, fill=vb, col=ifelse(bandType=="Trained","black",NA),size=ifelse(bandType=="Trained","black",NA))) + 
#   stat_bar + 
#     scale_color_manual(values = c("black" = "black"), guide = FALSE) +
#   scale_size_manual(values = c("black" = .5), guide = FALSE) +
#     ggh4x::facet_nested_wrap(~Resp+rename_fm(Fit_Method), axes = "all",ncol=3,scale="free") + 
#   theme(legend.position = "none",plot.title = element_text(hjust=.55) ) + labs(title="Model Predictions", y="vx", x="Condition")

pmod1 <- post_dat_l |> filter(!(Resp=="Observed")) |> 
  group_by(id,condit, Fit_Method,Resp,x) |> 
 summarize(vx=median(val)) |> 
 left_join(testAvgE1, by=join_by(id,condit,x==bandInt)) |>
 ggplot( aes(x=condit,y=vx, fill=vb,col=ifelse(bandType=="Trained","black",NA),size=ifelse(bandType=="Trained","black",NA))) + 
  stat_bar + 
    facet_wrap(~Resp+rename_fm(Fit_Method), strip.position = "top", scales = "free_x") +
        scale_color_manual(values = c("black" = "black"), guide = FALSE) +
  scale_size_manual(values = c("black" = .5), guide = FALSE) +
    theme(panel.spacing = unit(0, "lines"), 
         strip.background = element_blank(),
         strip.placement = "outside",
         legend.position = "none",plot.title = element_text(hjust=.50),
         axis.title.x = element_blank(),
         plot.margin = unit(c(20,0,0,0), "pt")) + 
         labs(title="Model Predictions - Experiment 1 Data", y="vx")

(pemp1)  / pmod1 + 
  plot_layout(design = layout) #heights = unit(c(5,-5, 8), c('cm','null'))

@@

```{r}
#| label: fig-cm-dev-pat
#| fig-cap: "Empirical data and Model predictions for absolute deviation from target (accuracy). Observed data is shown on top, and predictions from each combination of model and fitting method are shown below. "
#| fig-height: 11
#| fig-width: 11

pemp1 <- e1 |> filter(expMode2=="Test") |> 
  group_by(id,condit,vb,bandType) |> 
  ggplot(aes(x=condit,y=dist, fill=vb, col=ifelse(bandType=="Trained","black",NA),size=ifelse(bandType=="Trained","black",NA))) + 
  stat_summary(fun=mean, geom="bar", position=position_dodge()) + 
  stat_summary(fun.data=mean_se, geom="errorbar", color="black", position=position_dodge(), size=.5) + 
  scale_color_manual(values = c("black" = "black"), guide = FALSE) +
  scale_size_manual(values = c("black" = .5), guide = FALSE) + 
  scale_y_continuous( breaks=seq(0,600,by=200),labels=as.character(seq(0,600,by=200))) +
  expand_limits(y=600) +
  theme(legend.position="right", axis.title.x = element_blank(),plot.title = element_text(hjust=.40)) +
  labs(title="Empirical Data", y="Deviation from target band", fill="Band")

# layout <- "#A#
# #B#
# CCC"

layout <- "#A#
CCC"

# pmod1 <- post_dat_l |> filter(!(Resp=="Observed")) |> 
#   left_join(testAvgE1, by=join_by(id,condit,x==bandInt)) |>
#   group_by(id,condit, Fit_Method,Resp,x,bandType) |> 
#   mutate(x=as.factor(x)) |>
#  summarize(dist=median(dist)) |> 
#   mutate(oc = ifelse(bandType == "Trained", "black", NA)) |> 
#   ggplot(aes(x=condit,y=dist, fill=x,col=oc,size=oc)) + 
#   stat_bar +
#   ggh4x::facet_nested_wrap(~Resp+rename_fm(Fit_Method), axes = "all",ncol=3) + 
#    scale_color_manual(values = c("black" = "black"), guide = FALSE) +
#    scale_size_manual(values = c("black" = .3), guide = FALSE) + 
#   theme(legend.position = "none",plot.title = element_text(hjust=.55) ) + 
#   labs(title="Model Predictions", y="Deviation from target band", x="Condition")

pmod1 <- post_dat_l |> filter(!(Resp=="Observed")) |> 
  left_join(testAvgE1, by=join_by(id,condit,x==bandInt)) |>
  group_by(id,condit, Fit_Method,Resp,vb,bandType) |> 
 summarize(dist=median(dist)) |> 
 ggplot( aes(x=condit,y=dist, fill=vb,col=ifelse(bandType=="Trained","black",NA),size=ifelse(bandType=="Trained","black",NA))) + 
  stat_bar + 
    facet_wrap(~Resp+rename_fm(Fit_Method), strip.position = "top", scales = "free_x") +
        scale_color_manual(values = c("black" = "black"), guide = FALSE) +
  scale_size_manual(values = c("black" = .5), guide = FALSE) +
    theme(panel.spacing = unit(0, "lines"), 
         strip.background = element_blank(),
         strip.placement = "outside",
         legend.position = "none",
         plot.title = element_text(hjust=.50),
         plot.margin = unit(c(20,0,0,0), "pt"),
         axis.title.x = element_blank()) + 
         labs(title="Model Predictions - Experiment 1 Data", y="Deviation from target band")

(pemp1) / pmod1 + plot_layout(design = layout) #heights = unit(c(5,-5, 8), c('cm','null'))

@@

testAvgE1 |> ggplot( aes(x=(interaction(vb,bandType,sep="!")),y=vxAvg, fill=vb)) + 
    stat_summary(fun=mean, geom="bar", position="dodge", alpha=.75)+
    facet_wrap(~condit, scales = "free_x") +
    theme( ggh4x.axis.nestline  = element_line(colour = "black"),
           strip.background = element_blank(),
           strip.placement = "outside",
            axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5),) +   
    scale_x_discrete(guide = guide_axis_nested(delim = "!"), name = "Condit")

# stat_bar <- list(stat_summary(fun=mean, geom="bar", position=position_dodge(), alpha=.75),
#   stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge()))
# 
#   scale_x_discrete(limits = c(levels(df_plot$Top_contributor)[1:4],
#                               "ABC",
#                               levels(df_plot$Top_contributor)[5:8],
#                               "DEF",
#                               levels(df_plot$Top_contributor)[9:12]),
#                    labels = c("ABC" = "",
#                               "DEF" = ""))
```

@@

e3_model <- load_e3()

### 2024-03-21 Model/htw_model_e1.qmd
tweaking

>>>>

post_tabs$et_sum

et_sum <- post_dat_l |>
  group_by(id,condit, Fit_Method, Resp) |>
  summarise(val = mean(val), .groups = 'drop') |>
  pivot_wider(
    names_from = Resp,
    values_from = val,
    values_fill = list(val = NA)
  ) |>
  mutate(
    ALM_error = round(abs(ALM - Observed),1),
    EXAM_error = round(abs(EXAM - Observed),1),
    Best_Model = case_when(
      ALM_error < EXAM_error ~ "ALM",
      EXAM_error < ALM_error ~ "EXAM",
      TRUE ~ NA_character_  # In case of a tie or missing data
    )
  ) |>
  group_by(condit, Fit_Method) %>%
  summarise(
    Avg_ALM_error = round(mean(ALM_error, na.rm = TRUE),1),
    Avg_EXAM_error = round(mean(EXAM_error, na.rm = TRUE),1),
    N_Best_ALM = sum(Best_Model == "ALM", na.rm = TRUE),
    N_Best_EXAM = sum(Best_Model == "EXAM", na.rm = TRUE)
  ) %>%
  mutate(
    Best_Model = case_when(
      Avg_ALM_error < Avg_EXAM_error ~ "ALM",
      Avg_EXAM_error < Avg_ALM_error ~ "EXAM",
      TRUE ~ "Tie"  # In case of a tie or missing data
    ))

et_sum_x <- post_dat_l |>
  group_by(id,condit, Fit_Method, Resp,x) |>
  summarise(val = mean(val), .groups = 'drop') |>
  pivot_wider(
    names_from = Resp,
    values_from = val,
    values_fill = list(val = NA)
  ) |>
  mutate(
    ALM_error = round(abs(ALM - Observed),1),
    EXAM_error = round(abs(EXAM - Observed),1),
    Best_Model = case_when(
      ALM_error < EXAM_error ~ "ALM",
      EXAM_error < ALM_error ~ "EXAM",
      TRUE ~ NA_character_  # In case of a tie or missing data
    )
  ) |>
  group_by(condit, Fit_Method) %>%
  summarise(
    Avg_ALM_error = round(mean(ALM_error, na.rm = TRUE),1),
    Avg_EXAM_error = round(mean(EXAM_error, na.rm = TRUE),1),
    N_Best_ALM = sum(Best_Model == "ALM", na.rm = TRUE),
    N_Best_EXAM = sum(Best_Model == "EXAM", na.rm = TRUE)
  ) %>%
  mutate(
    Best_Model = case_when(
      Avg_ALM_error < Avg_EXAM_error ~ "ALM",
      Avg_EXAM_error < Avg_ALM_error ~ "EXAM",
      TRUE ~ "Tie"  # In case of a tie or missing data
    ))

et_sum_x_id <- post_dat_l |>
  group_by(id,condit, Fit_Method, Resp,x) |>
  summarise(val = mean(val), .groups = 'drop') |>
  pivot_wider(
    names_from = Resp,
    values_from = val,
    values_fill = list(val = NA)
  ) |>
  mutate(
    ALM_error = round(abs(ALM - Observed),1),
    EXAM_error = round(abs(EXAM - Observed),1),
    Best_Model = case_when(
      ALM_error < EXAM_error ~ "ALM",
      EXAM_error < ALM_error ~ "EXAM",
      TRUE ~ NA_character_  # In case of a tie or missing data
    )
  ) |>
  group_by(condit, Fit_Method) %>%
  summarise(
    Avg_ALM_error = round(mean(ALM_error, na.rm = TRUE),1),
    Avg_EXAM_error = round(mean(EXAM_error, na.rm = TRUE),1),
    N_Best_ALM = sum(Best_Model == "ALM", na.rm = TRUE),
    N_Best_EXAM = sum(Best_Model == "EXAM", na.rm = TRUE)
  ) %>%
  mutate(
    Best_Model = case_when(
      Avg_ALM_error < Avg_EXAM_error ~ "ALM",
      Avg_EXAM_error < Avg_ALM_error ~ "EXAM",
      TRUE ~ "Tie"  # In case of a tie or missing data
    ))

 agg_pred_full <-  post_dat  |> group_by(id,condit,Model,Fit_Method,x) |> 
    mutate(e2=abs(y-pred)) |> 
    summarise(y=mean(y), pred=mean(pred), mean_error=mean(e2),ei2=abs(y-pred)) |>
    group_by(id,condit,Model,Fit_Method) |> 
    summarise(mean_error=mean(mean_error),imean_error=mean(ei2)) |> 
    round_tibble(1) |> 
    group_by(Fit_Method,Model,condit) |> 
   # summarise(mean_error=mean(mean_error),imean_error=mean(imean_error))
   summarise(mean_error=mean(imean_error))

   agg_pred_full <-  pd_train  |> group_by(condit,Model,Fit_Method,Block,x) |> 
    mutate(e2=abs(y-pred)) |> 
    summarise(y=mean(y), pred=mean(pred), mean_error=mean(e2),ei2=abs(y-pred)) |>
    group_by(condit,Model,Fit_Method,Block) |> 
    summarise(mean_error=mean(mean_error), mean_ei2=mean(ei2)) |> 
    group_by(Fit_Method,Model,condit) |>
    summarise(mean_error=mean(mean_error),mean_ei2=mean(mean_ei2))

 pd_train  |> group_by(id, condit,Model,Fit_Method,Block,x) |> 
    mutate(e2=abs(y-pred)) |> 
    summarise(y=mean(y), pred=mean(pred), mean_error=mean(e2),ei2=abs(y-pred)) |>
    group_by(condit,Model,Fit_Method,Block) |> 
    summarise(mean_error=mean(mean_error), mean_ei2=mean(ei2)) |> 
    group_by(Fit_Method,Model,condit) |>
    summarise(mean_error=mean(mean_error),mean_ei2=mean(mean_ei2))
